---
title: "VARIANCE INFLATION FACTOR AS A TEST OF MULTICOLINEARITY IN MULTIPLE REGRESSION MODEL"
author: "Daniel James"
date: "`r Sys.Date()`"
output: html_document
---
  
  ```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "",
                      collapse = TRUE,
                      echo = TRUE,
                      error = TRUE, # do not interrupt in case of errors
                      message = FALSE,
                      warning = FALSE,
                      comma <- function(x) format(x, digits = 2, big.mark = ",")
)
````
### Load packages and prepare multicore process
### This recipe explains how to check multicollinearity in regression using R.

### Variance inflation factor (VIF) is used for detecting the multicollinearity in a model, which measures the correlation and strength of correlation between the independent variables in a regression model. - If the value of VIF is less than 1: no correlation - If the value of VIF is between 1-5, there is moderate correlation - If the value of VIF is above 5: severe correlation

### Step 1 - Install necessary packages

```{r install_packages, include = TRUE}
library(pacman)

p_load(caTools)    # For Linear regression 
p_load(car)        # To check multicollinearity 
p_load(quantmod)
p_load(MASS)
p_load(corrplot)   # plot correlation plot
````

### Step 2 - Define a Dataframe

```{r creat_data_frame, include = TRUE}
data <- data.frame(marks_scored = c(35,42,24,27,37),        # marks : dependent variable(y)
                   no_hours_studied = c(5,4,2,3,4),
                   no_hours_played = c(4,3,4,2,2),
                   attendance = c(8,8,4,6,9))

head(data)
````

### Step 3 - Create a linear regression model

```{r Create_a_linear_regression_model, include = TRUE}
model_all <- lm(marks_scored ~ ., data=data)  # with all the independent variables in the dataframe

summary(model_all)
````

### Step 4 - Use the vif() function

```{r Use_the_vif_function, include = TRUE}
vif(model_all)
````

### Step 5 - Visualize VIF Values

```{r Visualize_VIF_Values, include = TRUE}

vif_values <- vif(model_all)           #create vector of VIF values

barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue") #create horizontal bar chart to display each VIF value

abline(v = 5, lwd = 3, lty = 2)    #add vertical line at 5 as after 5 there is severe correlation
````

### After plotting the graph, user can does decide which variable to remove i.e not include in model building and check whether the coreesponding R squared value improves.

### Step 6 - Multicollinearity test can be checked by

```{r check_for_multicolinearity, include = TRUE}
data_x <- data[,2:4]                                       # independent variables 

var <- cor(data_x)                                         # independent variables correlation matrix 

var_inv <- ginv(var)                                       # independent variables inverse correlation matrix 

colnames(var_inv) <- colnames(data_x)                      # rename the row names and column names
rownames(var_inv) <- colnames(data_x)

corrplot(var_inv,method='number',is.corr = F)              # visualize the multicollinearity
#{"mode":"full","isActive":false}
````
